Single-cell virus sequencing of influenza infections that trigger innate immunity

### 2018-10-04 paper/paper.tex
added Jacob undergrad grants
added a bit of text to Discussion

\section{Introduction}
Infection with an acute virus such as influenza initiates a race between the virus and immune system.
As the virus spreads, some infected cells trigger innate immunity and begin producing interferon (IFN).
This IFN directs expression of anti-viral interferon-stimulated genes (ISGs) in the infected cell and its neighbors via autocrine and paracrine signaling, as well as helping launch a broader immune response~\citep{stetson2006type,honda2006type}.
If innate immunity is activated sufficiently rapidly, it can reduce viral replication and disease severity~\citep{solov1969results,treanor1987intranasally,beilharz2007protection,kugel2009intranasal,steel2010transmission}---although excessive prolonged immune responses later in infection can actually be associated with immunopathology and more severe disease~\citep{la2007question, iwasaki2014innate}.

@@

However, existing techniques are inadequate to determine how viral genetic diversity contributes to cell-to-cell heterogeneity during infection, since they do not determine full sequences of the virions that infect individual cells.
Flow cytometry and fluorescent reporters only measure protein levels~\citep{brooke2013most, guo2017single}, and current single-cell transcriptomic techniques primarily measure abundance of transcripts~\citep{russell2018extreme, zanini2018single, zanini2018virus, steuerman2018dissection, saikia2018simultaneous}.
None of these techniques reliably reveal if the virion infecting a specific cell has some idiosyncratic mutation, despite the fact that such mutations are common regardless of whether the cells are infected with a nominally ``wildtype'' or ``mutant'' stock of virus.

@@

\clearpage
\section{Results}

@@

\section{Discussion}

\jdbcomment{Section needs to be complete.
Add this:
Approach builds on observations by \citet{saikia2018simultaneous} and \citet{zanini2018virus} that fragmentary viral genetic information can be obtained using more standard single-cell techniques, and is similar the approach used by \cite{gupta2018single} to obtain full-length isoforms of some cellular genes in single cells.
}

@@

\section{Acknowledgments}
We thank Cole Trapnell, Jason Underwood, Robert Bradley, Daniel Stetson, and AJ Velthuis for helpful scientific suggestions.
We thank Andy Marty and the Fred Hutch Genomics Core for performing the deep sequencing.
We thank Katherine Xue, David Bacsik, and Hannah Itell for comments on the manuscript.
This work was supported by the NIAID of the NIH under grant R01 AI127893 and a Burroughs Wellcome Fund Young Investigator in the Pathogenesis of Infectious Diseases grant to JDB.
ABR was supported by a postdoctoral fellowship from the Damon Runyon Cancer Research Foundation.
JDB is an Investigator of the Howard Hughes Medical Institute.
The funders had no role in study design, data collection and interpretation, or the decision to submit the work for publication.
The authors declare that no competing interests exist.

>>>>

\section{Introduction}
Infection with an acute virus such as influenza initiates a race between the virus and immune system.
As the virus spreads, some infected cells trigger innate immunity and begin producing interferon (IFN).
This IFN directs expression of anti-viral interferon-stimulated genes (ISGs) in the infected cell and its neighbors via autocrine and paracrine signaling, as well as helping launch a broader immune response~\citep{stetson2006type,honda2006type}.
If innate immunity is activated sufficiently rapidly, it can reduce viral replication and disease severity~\citep{solov1969results,treanor1987intranasally,beilharz2007protection,kugel2009intranasal,steel2010transmission}---although excessive immune responses later in infection can actually be associated with immunopathology and more severe disease~\citep{la2007question, iwasaki2014innate}.

@@

However, existing techniques are inadequate to determine how viral genetic diversity contributes to cell-to-cell heterogeneity during infection, since they do not determine full sequences of the virions that infect individual cells.
Flow cytometry and fluorescent reporters only measure protein levels~\citep{brooke2013most, guo2017single}, and current single-cell transcriptomic techniques primarily measure abundance of transcripts and provide only fragmentary information on their sequences~\citep{russell2018extreme, zanini2018single, zanini2018virus, steuerman2018dissection, saikia2018simultaneous}.
None of these techniques reliably reveal if the virion infecting a specific cell has some idiosyncratic mutation, despite the fact that such mutations are common regardless of whether the cells are infected with a nominally ``wildtype'' or ``mutant'' stock of virus.

@@

\section{Results}

@@

\section{Discussion}
We have determined the full sequences of the infecting virions in single influenza-infected cells, and then examined how viral mutations and gene-expression defects associate with infection outcome in each cell.
Methodologically, the major advance of our work is to measure the \emph{genotypes} of viruses in single cells in addition to quantifying the abundance of viral components (i.e., transcripts, proteins, or progeny virions) as has been done by prior single-cell virology studies~\citep{russell2018extreme, zanini2018single, zanini2018virus, steuerman2018dissection, saikia2018simultaneous, zhu2009growth, schulte2014single, akpinar2016high, heldt2015single, brooke2013most}.
Our method builds on the observations by \citet{saikia2018simultaneous} and \citet{zanini2018virus} that fragmentary viral genetic information can be obtained using standard single-cell transcriptomic techniques.
To make this genetic information complete, we have coupled a standard single-cell transcriptomic technique to long-read PacBio sequencing, a strategy analogous to that recently used by \cite{gupta2018single} to obtain full-length isoforms of some cellular genes in single cells.

Our results demonstrate that the single-cell viral genetic information provided by this methodological advance is extremely valuable for understanding the biology of influenza infection, especially as it relates to innate-immune detection.

@@

\section{Acknowledgments}
We thank Cole Trapnell, Jason Underwood, Robert Bradley, Daniel Stetson, and AJ Velthuis for helpful scientific suggestions.
We thank Andy Marty and the Fred Hutch Genomics Core for performing the deep sequencing.
We thank Katherine Xue, David Bacsik, and Hannah Itell for comments on the manuscript.
This work was supported by the NIAID of the NIH under grant R01 AI127893 and a Burroughs Wellcome Fund Young Investigator in the Pathogenesis of Infectious Diseases grant to JDB.
ABR was supported by a postdoctoral fellowship from the Damon Runyon Cancer Research Foundation.
JRK was supported by a Washington Research Foundation Undergraduate Research Fellowship and a Mary Gates undergraduate research scholarship from the University of Washington.
JDB is an Investigator of the Howard Hughes Medical Institute.
The funders had no role in study design, data collection and interpretation, or the decision to submit the work for publication.
The authors declare that no competing interests exist.
